Post-replication citation patterns in psychology: Four case studies

### 2021-02-07 manuscript/manuscript.Rmd
finalizing alignment of google doc and r markdown

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, University of Amsterdam"
  - id            : "2"
    institution   : "Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for Transforming Biomedical Research, Charité – Universitätsmedizin Berlin"
  - id            : "3"
    institution   : "Department of Psychology, University of Cambridge, UK"
  - id            : "4"
    institution   : "School of Psychological Science, University of Bristol"
  - id            : "5"
    institution   : "MRC Integrative Epidemiology Unit at the University of Bristol"
  - id            : "6"
    institution   : "Department of Methodology and Statistics, Tilburg School of Social and Behavioral Sciences, Tilburg University"
  - id            : "7"
    institution   : "Meta-Research Innovation Center at Stanford (METRICS), Stanford University"
  - id            : "8"
    institution   : "Departments of Medicine, of Health Research and Policy, of Biomedical Data Science, and of Statistics, Stanford University"

@@

In total, `r d_citations %>% nrow()` articles cited one of the original studies of which `r d_contentAnalysis %>% nrow()` articles (after taking a 40% random sample in the Baumeister case) fell within the time period designated for qualitative assessment. Of these `r d_contentAnalysis %>% nrow()` articles, we excluded `r excluded_n` from the qualitative analysis because (1) we could not access the full text (n = `r noAccess_n`); (2) they were non-English language (n = `r nonEnglish_n`); (3) they included a citation to the original study in the reference section, but not in the main text (n = `r refCiteOnly_n`); or (4) manual inspection indicated that they did not actually appear to cite the original study at all (n = `r noCite_n`). Research design classifications for the remaining `r notexcluded_n` articles included in the qualitative analysis are shown in Table \@ref(tab:articleTypes).

@@

```{r articleTypes}
d_contentAnalysis %>%
  filter(excluded == F) %>%
  count(articleType) %>%
  mutate("Count (%)" = paste0(n, " (", round(n / sum(n), 2) * 100, ")")) %>%
  select("Article type" = articleType, "Count (%)", -n) %>%
  apa_table(caption = "Counts and percentages for research design classifications of articles included in qualitative analyses")
```

@@

>>>>

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, University of Amsterdam"
  - id            : "2"
    institution   : "Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for Transforming Biomedical Research, Charité – Universitätsmedizin Berlin"
  - id            : "3"
    institution   : "Department of Psychology, University of Cambridge, UK"
  - id            : "4"
    institution   : "School of Psychological Science, University of Bristol"
  - id            : "5"
    institution   : "MRC Integrative Epidemiology Unit at the University of Bristol"
  - id            : "6"
    institution   : "Department of Methodology and Statistics, Tilburg School of Social and Behavioral Sciences, Tilburg University"
  - id            : "7"
    institution   : "Meta-Research Innovation Center at Stanford (METRICS), Stanford University"
  - id            : "8"
    institution   : "Departments of Medicine, of Epidemiology and Population Health, of Biomedical Data Science, and of Statistics, Stanford University"

@@

In total, `r d_citations %>% nrow()` articles cited one of the original studies of which `r d_contentAnalysis %>% nrow()` articles (after taking a 40% random sample in the Baumeister case) fell within the time period designated for qualitative assessment. Of these `r d_contentAnalysis %>% nrow()` articles, we excluded `r excluded_n` from the qualitative analysis because (1) we could not access the full text (n = `r noAccess_n`); (2) they were non-English language (n = `r nonEnglish_n`); (3) they included a citation to the original study in the reference section, but not in the main text (n = `r refCiteOnly_n`); or (4) manual inspection indicated that they did not actually appear to cite the original study at all (n = `r noCite_n`). Article type classifications for the remaining `r notexcluded_n` articles included in the qualitative analysis are shown in Table \@ref(tab:articleTypes).

@@

```{r articleTypes}
d_contentAnalysis %>%
  filter(excluded == F) %>%
  count(articleType) %>%
  mutate("Count (%)" = paste0(n, " (", round(n / sum(n), 2) * 100, ")")) %>%
  select("Article type" = articleType, "Count (%)", -n) %>%
  apa_table(caption = "Counts and percentages for article type classifications of articles included in qualitative analyses")
```

@@

The findings presented here are inherently limited by the observational nature of the study design, which complicates straightforward conclusions about the causal impact of the replications. Although the use of a reference class enables us to detect the influence of exogenous factors to some extent, we cannot rule out their contribution. For example, the modest decline in favourable citations observed in most cases could be attributable, at least in part, to a more general awareness in the research community about methodological issues (for example, that the sample sizes of the original studies may not have provided adequate statistical power). We have also focused only on the replication study and the original study in each case study, without considering the impact of other potentially relevant events. Notably, meta-research studies have detected signatures of publication bias and other questionable research practices in the fields to which these case studies belong (E. C. Carter et al., 2015; Coles et al., 2019; Vadillo, 2019; Vadillo et al., 2016) and other relevant replication studies contesting prior findings have been published (e.g., Rohrer et al., 2015).
